---
title: "Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
library(ggplot2)
```


# Reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = F)
```

* calculate annual escapment by species and region
  + `separate`
  + `group_by`
  + `summarize` using the `sum` function
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)
  + `filter`

annual_esc

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% # coerce year to numeric type
  group_by(year, Species, SASAP.Region) %>% 
  summarise(count = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))
```


# Generate some static plots

```{r}
ggplot(annual_esc, mapping = aes(x = Species, y = count, fill = SASAP.Region)) +
  geom_col()
```

```{r}
my_theme <- theme_test() +
  theme(legend.position = "bottom", legend.title = element_blank())
```


```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement (num of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```


# Generate an interactive map