---
title: "Data Visualization"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
library(DT)
```


# Reformat the data for plotting

For dyplr/tidyr reference see my [data tidying page](data-tidying.html)

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M. [https://doi.org/10.5063/F1S46Q6M](https://doi.org/10.5063/F1S46Q6M)

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = F)
```

* calculate annual escapment by species and region
    + `separate`
    + `group_by`
    + `summarize` using the `sum` function
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)
    + `filter`

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% # coerce year to numeric type
  group_by(year, Species, SASAP.Region) %>% 
  summarise(count = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))
```


# Generate some static plots

```{r, echo = F}
ggplot(annual_esc, mapping = aes(x = Species, y = count, fill = SASAP.Region)) +
  geom_col()
```

```{r, echo = F}
my_theme <- theme_test() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))
```


```{r, echo = F}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) +
  ylab("Escapement (num of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```

```{r, fig.width = 7, fig.height = 8, echo = F}
p <- ggplot(annual_esc, aes(x = year, y = count, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(label = comma) +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
  my_theme
p

ggsave("figures/regional_escapement.png", 
       plot = p,
       width = 7,
       height = 8,
       units = "in")
```


# Generate an interactive map

Map the sampling locations for escapement

<br/>
<br/>
<br/>

* find distinct lat/lon/location combinations

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
```

```{r, echo = F}
datatable(locations)
```

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~ Location)
```

```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```

