---
title: "Untitled"
author: "Jeanette Clark"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, warning = F, message = F, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)
```

# Data Tidying

Escapement data are source from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r, echo = F, results = F}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), 
                stringsAsFactors = F)

head(esc)
```

Calculate annual, regional total escapement by species

* `separate` the date into year - month - day
* `group_by` `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species

```{r, echo = F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

```

```{r, echo = F}
datatable(annual_esc)
```


# Static Plots
```{r}
my_theme <- theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(annual_esc, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col() +
  my_theme
```

timeseries of Kodiak escapement by species

```{r}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
       aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab('Escapement (number of fish)') +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
  
```

Plot timeseries of escapement by species for all regions

```{r, fig.height = 10, fig.width=8}
plot_name <- ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free", ncol = 2) +
  my_theme

plot_name

ggsave("figures/region_escapement.png", 
       plot = plot_name, 
       height = 10,
       width = 8,
       units = "in")
```

```{r}
write.csv(annual_esc, 
          "derived_data/annual_escapement.csv", 
          row.names = F)
```

# Interactive Map

Make a map using leaflet

* find unique locations with lat/lon values

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

datatable(locations)
```

Create our interactive map

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

